---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```
# splicr: multiple argument piping in R

<!-- badges: start -->
[![Travis build status](https://travis-ci.org/Christopher-Peterson/splicr.svg?branch=master)](https://travis-ci.org/Christopher-Peterson/splicr)
<!-- badges: end -->

The `splicr` package implements a new pipe operator, `%!>%`, which uses the values of a list on the left as arguments for a right-hand function call.

```{r example1}
library(splicr)
example_fun = function(a, b = "B", c = "C", d = "D") {
  paste0("a = ", a, ", b = ", b, ", c = ", c, ", d = ", d)
}
list("A", c = "Something Else") %!>% example_fun(d = "Alt D Value")
```

## Installation

You can install the development version of `splicr` with
```{r installgh, eval = FALSE}
remotes::install_github("Christopher-Peterson/splicr")
```

## Splicing rules

The multi-pipe follows a few simple rules for splicing arguments into a function:

### You can pipe into functions or function calls

The following lines all produce the same output:
```{r func_vs_call}
list(a = "A", b = "B") %!>% example_fun
list(a = "A", b = "B") %!>% example_fun()
```

Since no arguments were specified on the right hand side, these are equivalent to `do.call()` and similar alternatives.

```{r alternatives}
do.call(example_fun, list(a = "A", b = "B"))
rlang::exec(example_fun, a = "A", b = "B")
rlang::exec(example_fun, !!!list(a = "A", b = "B"))
```

However, `splicr`'s partial argument specification provides greater flexibility than `do.call()`, and using `magrittr::%>%` to pipe  into `rlang::exec` with the splice operator (`!!!`) with has a rather convoluted syntax:
```{r magrittr_splice}
library(magrittr)
list(a = "A", b = "B") %>% {rlang::exec(.fn = example_fun, !!!.)}
```

### Positional arguments have left precedence 

If there are unnamed arguments on both the left and right, the left arguments will go first.

```{r left_prec}
list("A", "B") %!>% example_fun("Not A", "Not B")
```

Unnamed arguments will be placed in the first available position.

```{r left_prec2}
list("A", "Not B") %!>% example_fun(b = "B", "Not C")
```

### Named arguments have right precedence

If there same argument is named on both sides, the right hand version will be used (with a warning, by default).

```{r right_prec}
list("A", b = "Not B") %!>% example_fun(b = "B", "Not C")
```

### Annonymous functions are supported, if wrapped in braces

```{r annon1}
list(a = 1, b = 2) %!>% { function(a, b) a + b } 
```

`purrr`-style lambda functions are also supported:

```{r annon2}
list(1,  2, 3) %!>% { ~ .x + .y + ..3} 
```

Not wrapping lambdas or inline functions in braces this causes strange parser errors. The multi-pipe tries to detect when this happens and fail with a useful error message. 

### Wrap function factories in parentheses to pipe to their outputs

Piping to a function factory (a function that returns other functions) will normally splice the arguments into the factory.

```{r factory}
factory = function(a = "A") {
  function(b, c = "C")  paste(a, b, c)
}
out <- list("Hrm?") %!>% factory
# out is a function
list("B", "C") %!>% out
```

To splice into the factory's output, wrap the right hand side in parentheses.

```{r factory2}
list("B") %!>% (factory("Hrm?"))
```

## Splice pipes are compatable with `magrittr::%>%`

It's straightforward to combine `%!>%` with `%>%` as part of chain:

```{r with_mag} 
library(purrr, quietly = TRUE)

# simulate some random numbers, and see how much
# their range differs among samples
set.seed(2341)
list(.x = rpois(10, lambda = 5), # .x is count
     .y =  50 * rexp(10)) %!>%    # .y is scale
  map2(~runif(n = .x, min = -.y, max = .y)) %>% # random number set
  map(range) %>% # range of each set
  transpose %!>% # convert to a list of min and max
  map2_dbl(~.y - .x)

```

You can also use the pipe to easily access S3 object elements:

```{r lm_plot, results = FALSE, out.width='60%'}
lm(Sepal.Length ~ Sepal.Width, data = iris) %!>% {
    function(fitted.values, residuals, ...){
      plot(fitted.values, residuals)
      invisible(list(...))
    }
}
```
